\LARGE \bf Energy-Aware Dynamic Mission Planning Algorithm for UAVs

### 2020-11-23 icra-2021.tex
fixes to the lemma but still more fixes to come

%%
%% Conference Paper for ICRA'21, May 16-22, Xi'an, China
%% 
%%

@@

Let us consider a Fourier series of an arbitrary order $r\in\mathbb{Z}_{\geq 0}$ and period $T\in\mathbb{Z}_{> 0}$ for the purpose of energy consumption modeling of the mission
\begin{equation}\label{eq:fourier}
  h(k)=a_0/T+(2/T)\sum_{j=1}^{r}{\left(a_j\cos{\omega jk}+b_j\sin{\omega jk}\right)},
\end{equation}
where $h:\mathbb{R}_{\geq 0}\rightarrow\mathbb{R}$ maps time to the modeled instantaneous energy consumption, $\omega:=2\pi/T$ is the angular frequency, and $a,b\in\mathbb{R}$ the Fourier series coefficients.

Consider a linear time-invariant (LTI) state-space model
\begin{equation}\label{eq:state-perf}\begin{split}
  \mathbf{q}_{k+1}&=A\mathbf{q}_{k}+B\mathbf{u}_{k}+\mathbf{w}_k,\\
  y_k&=C\mathbf{q}_k+v_k,
\end{split}\end{equation}
where $y_k\in\mathbb{R}_{\geq 0}$ is the instantaneous energy consumption. The state $\mathbf{q}$ are the energy coefficients
\begin{equation}\label{eq:state-details}\begin{split}
  \mathbf{q}_k&=\left[\begin{array}{cccccc}
    \alpha_0 & \alpha_1 & \beta_1 & \cdots & \alpha_r & \beta_r
  \end{array}\right]^T,\\
  A&=\left[\begin{array}{cccc}
    1&    &       &  \\
     & A_1&       &  \\
     &    & \ddots&  \\
     &    &       & A_r 
  \end{array}\right],\,A_j\begin{bmatrix}1 & \omega j \\ -\omega j & 1\end{bmatrix},\\
  C&=(1/T)\left[\begin{array}{cccccc}
    1 & \cdots & 1
  \end{array}\right],
\end{split}\end{equation}
where $\mathbf{q}_k\in\mathbb{R}^m$ given $m=2r+1$, $A\in\mathbb{R}^{m\times m}$ is the state transmission matrix, and $C\in\mathbb{R}^m$ is the output matrix. In matrix $A$, the top left entry is one, the diagonal entries are $A_1,\dots,A_r$, the remaining entries are zero.

\begin{lem}[Fourier series, LTI state-space equivalence]Suppose there is no state and output uncertainty (the contributions $\mathbf{w}_k\in\mathbb{R}^m,v_k\in\mathbb{R}$ are zero) and the control is zero. The non-linear model in Equation~(\ref{eq:fourier}) can be expressed using the LTI state-space model in Equation~(\ref{eq:state-perf})
\end{lem}
\begin{proof}
We re-write the Fourier series expression in Equation~(\ref{eq:fourier}) in its complex form with the well-known Euler's formula $e^{it}=\cos{t}+i\sin{t}$. With $t=\omega jk$, we find the expression for $\cos{\omega jk}=(e^{i\omega jk}+e^{-i\omega jk})/2$ and $\sin{\omega jk}=(e^{i\omega jk}-e^{-i\omega jk})/(2i)$ by substitution of $\sin{\omega jk}$ and $\cos{\omega jk}$ respectively. This leads~\cite{kuo1967automatic}
\begin{equation}\begin{split}\label{eq:proof-complex}
  h(k)=a_0/T+&(1/T)\sum_{j=1}^{r}{e^{i\omega jk}(a_j-ib_j)}+\\&(1/T)\sum_{j=1}^{r}{e^{-i\omega jk}(a_j+ib_j)},
 \end{split}\end{equation}
where $i$ is the imaginary unit.

Let us consider the continuous form of Equation~(\ref{eq:state-perf}), $\dot{\mathbf{q}}=A_c\mathbf{q}$, and an initial guess $\mathbf{q}_0$. $A_c$ is defined as the continuous version of matrix $A$ in Equation~(\ref{eq:state-details}); $A_c:=A-I$, where $I$ is a properly sized identity matrix. The solution at time $k$ can be expressed $\mathbf{q}=e^{A_ck}\mathbf{q}_0$. Both the solution and the system in Equation~(\ref{eq:state-perf}) are well established expressions derived using a standard textbooks~\cite{kuo1967automatic, ogata2002modern}. To solve the matrix exponential $e^{A_ck}$, we use the eigenvectors matrix decomposition method~\cite{moler2003nineteen}.

The method works on the similarity transformation $A_c=VDV^{-1}$. The power series definition of $e^{A_ck}$ implies $e^{A_ck}=Ve^{Dk}V^{-1}$~\cite{moler2003nineteen}. We consider the non-singular matrix $V$, whose columns are eigenvectors of $A_c$; $V:=\begin{bmatrix}v_0 & v_1^0 & v_1^1 & \dots & v_r^0 & v_r^1\end{bmatrix}$. We then consider the diagonal matrix of eigenvalues $D=\mathrm{diag}{(\lambda_0,\lambda_1^0,\lambda_1^1,\dots,\lambda_r^0,\lambda_r^1)}$. $\lambda_0$ is the eigenvalue associated to the first item of $A_c$. $\lambda_j^0,\lambda_j^1$ are the two eigenvalues associated with the block $A_j-I$. We can write $A_cv_j=\lambda_jv_j\,\,\,\forall j=\{1,\dots,m\}$, and $AV=VD$.

We apply the approach in terms of the continuous equivalent of the model in Equation~(\ref{eq:state-perf}), $\dot{\mathbf{q}}=A_c\mathbf{q}$. Initial guess is expressed as a combination of eigenvectors $\mathbf{q}_0=\mathbf{\gamma}V$. The generic solution is $\mathbf{q}_k=\mathbf{\gamma}e^{Dk}V$.

Let us express the output at time $k$ as a linear combination of the state $y=C\mathbf{q}=C\gamma e^{Dk}V$
\begin{equation}\label{eq:proof-output}\begin{split}
  y=\gamma_0e^{\lambda_0k}v_0/T+&(1/T)\sum_{j=1}^r{\gamma_je^{\lambda_j^0k}v_j^0}+\\&(1/T)\sum_{j=1}^r{\gamma_je^{\lambda_j^1k}v_j^1},
\end{split}\end{equation}
we proof that the eigenvalues $\mathbf{\lambda}$ and eigenvectors $V$ are such that Equation~(\ref{eq:proof-output}) is equivalent to Equation~(\ref{eq:proof-complex}).

The matrix $A_c$ is a block diagonal matrix, so we can express its determinant as the multiplication of the determinants of its blocks $\det{(A_c)}=\det{(0)}\times\det{(A_1-I)}\times\cdots\times\det{(A_r-I)}$. We proof each block separately.

First, we proof that the first term of both the Equation~(\ref{eq:proof-complex}) and~(\ref{eq:proof-output}) matches. We find the eigenvalue from $\det(0-\lambda I)=0$, which is $\lambda_0=0$. The corresponding eigenvector can be chosen arbitrarily $(0-\lambda_0)v_0=0\,\,\,\forall v_0$, thus we choose $v_0=1$. We find the value $\gamma_0$ of the vector $\gamma$ so that the terms are equal $a_0=\gamma_0$.

For the first block $A_1-I$, we find the eigenvalues from $\det(A_1-(1+\lambda)I)=0$. The polynomial $\lambda^2+\omega^2$, gives two complex roots--the two eigenvalues $\lambda_1^0=i\omega$ and $\lambda_1^1=-i\omega$. The eigenvector associated with the eigenvalue $\lambda_1^0$ is $v_1^0=\begin{bmatrix}-i&1\end{bmatrix}^T$. The eigenvector associated with the eigenvalue $\lambda_1^1$ is $v_1^1=\begin{bmatrix}i&1\end{bmatrix}^T$. Again, we find the subvector $\gamma_1$ of the vector $\gamma$ such that the equivalences $e^{i\omega k}(a_1-ib_1)=\gamma_1e^{i\omega k}\begin{bmatrix}-i&1\end{bmatrix}^T$ and $e^{-i\omega k}(a_1+ib_1)=\gamma_1e^{i\omega k}\begin{bmatrix}i&1\end{bmatrix}^T$ hold. They hold for $\gamma_1=\begin{bmatrix}b_1&a_1\end{bmatrix}$.

The discretized version $A$ of the matrix $A_c$ can be expressed $A=A_c+I$ for small enough values of time step (forward Euler approximation\footnote{One can guarantee to have the same outputs rather than an approximation with $A=e^{A_ck}$}). We thus conclude that the lemma holds.

\end{proof}

@@

\leavevmode\thispagestyle{empty}\newpage % just for now
\leavevmode\thispagestyle{empty}\newpage

@@

\appendices
\section{Mission Plan Format}

>>>>

%%
%% 
%% 
%%

@@

Let us consider a periodic signal of period $T\in\mathbb{Z}_{> 0}$ representing the energy consumption, and Fourier series of an arbitrary order $r\in\mathbb{Z}_{\geq 0}$ and for the purpose of energy consumption modeling of the signal
\begin{equation}\label{eq:fourier}
  h(t)=a_0/T+(2/T)\sum_{j=1}^{r}{\left(a_j\cos{\omega jt}+b_j\sin{\omega jt}\right)},
\end{equation}
where $h:\mathbb{R}_{\geq 0}\rightarrow\mathbb{R}$ maps time to the modeled instantaneous energy consumption, $\omega:=2\pi/T$ is the angular frequency, and $a,b\in\mathbb{R}$ the Fourier series coefficients.

Consider a linear time-invariant (LTI) state-space model
\begin{equation}\label{eq:state-perf}\begin{split}
  \dot{\mathbf{q}}&=A\mathbf{q}+B\mathbf{u}+\mathbf{w},\\
  y&=C\mathbf{q}+v,
\end{split}\end{equation}
where $y\in\mathbb{R}_{\geq 0}$ is the instantaneous energy consumption. The state $\mathbf{q}$ are the energy coefficients
\begin{equation}\label{eq:state-details}\begin{split}
  \mathbf{q}&=\left[\begin{array}{cccccc}
    \alpha_0 & \alpha_1 & \beta_1 & \cdots & \alpha_r & \beta_r
  \end{array}\right]^T,\\
  A&=\left[\begin{array}{cccc}
    0&    &       &  \\
     & A_1&       &  \\
     &    & \ddots&  \\
     &    &       & A_r 
  \end{array}\right],\,A_j\begin{bmatrix}0 & \omega j \\ -\omega j & 0\end{bmatrix},\\
  C&=(1/T)\left[\begin{array}{cccccc}
    1 & 1 & 0 &\cdots & 1 & 0
  \end{array}\right],
\end{split}\end{equation}
where $\mathbf{q}\in\mathbb{R}^m$ given $m=2r+1$, $A\in\mathbb{R}^{m\times m}$ is the state transmission matrix, and $C\in\mathbb{R}^m$ is the output matrix. In matrix $A$, the top left entry is one, the diagonal entries are $A_1,\dots,A_r$, the remaining entries are zero.

\begin{lem}[Representations equivalence]\label{lem:eqv}Suppose there is no state and output uncertainty (the contributions $\mathbf{w}\in\mathbb{R}^m,v\in\mathbb{R}$ are zero) and the control is zero. Suppose further matrices $A,C$ are described by Equation~(\ref{eq:state-details}), and the initial guess $\mathbf{q}_0$ is $\mathbf{q}_0=\begin{bmatrix}a_0 & a_1/2 & b_1/2 & \cdots & a_r/2 & b_r/2\end{bmatrix}$. Then Equation~(\ref{eq:fourier}) describes the same signal as Equation~(\ref{eq:state-perf}).
\end{lem}

The equivalence of the models is trivial and the coefficients are chosen such that the two Equations produce the same expression. We refer the reader to Appendix~\ref{app:proof-eqv} for a formal proof.

Let us consider the discretized version $A_d$ of the matrix $A$ can be expressed $A_d=A+I$ for small enough values of time step (forward Euler approximation\footnote{One can guarantee to have the same outputs rather than an approximation with $A=e^{A_ck}$}). % where to place this... Maybe in the algorithm...
% and write little more how, why you discretized...

@@

\leavevmode\thispagestyle{empty}\newpage % just for now

@@

\appendices

\section{Proof of Lemma~\ref{lem:eqv}}
\label{app:proof-eqv}

This appendix is a formal proof of Lemma~\ref{lem:eqv}.

The proof justifies the choice of the items of the matrices $A,C$ and of the initial guess $\mathbf{q}_0$ such that the coefficients $a_0,\dots,b_r$ are the same as the coefficients $\alpha_0,\dots,\beta_r$.

We re-write the Fourier series expression in Equation~(\ref{eq:fourier}) in its complex form with the well-known Euler's formula $e^{it}=\cos{t}+i\sin{t}$. With $t=\omega jt$, we find the expression for $\cos{\omega jt}=(e^{i\omega jt}+e^{-i\omega jt})/2$ and $\sin{\omega jt}=(e^{i\omega jt}-e^{-i\omega jt})/(2i)$ by substitution of $\sin{\omega jt}$ and $\cos{\omega jt}$ respectively. This leads~\cite{kuo1967automatic}
\begin{equation}\begin{split}\label{eq:proof-complex}
  h(t)=a_0/T+&(1/T)\sum_{j=1}^{r}{e^{i\omega jt}(a_j-ib_j)}+\\&(1/T)\sum_{j=1}^{r}{e^{-i\omega jt}(a_j+ib_j)},
 \end{split}\end{equation}
where $i$ is the imaginary unit.

The solution at time $t$ can be expressed $\mathbf{q}=e^{At}\mathbf{q}_0$. Both the solution and the system in Equation~(\ref{eq:state-perf}) are well established expressions derived using a standard textbooks~\cite{kuo1967automatic, ogata2002modern}. To solve the matrix exponential $e^{At}$, we use the eigenvectors matrix decomposition method~\cite{moler2003nineteen}.

The method works on the similarity transformation $A_c=VDV^{-1}$. The power series definition of $e^{At}$ implies $e^{At}=Ve^{Dt}V^{-1}$~\cite{moler2003nineteen}. We consider the non-singular matrix $V$, whose columns are eigenvectors of $A$; $V:=\begin{bmatrix}v_0 & v_1^0 & v_1^1 & \dots & v_r^0 & v_r^1\end{bmatrix}$. We then consider the diagonal matrix of eigenvalues $D=\mathrm{diag}{(\lambda_0,\lambda_1^0,\lambda_1^1,\dots,\lambda_r^0,\lambda_r^1)}$. $\lambda_0$ is the eigenvalue associated to the first item of $A$. $\lambda_j^0,\lambda_j^1$ are the two eigenvalues associated with the block $A_j$. We can write $A_cv_j=\lambda_jv_j\,\,\,\forall j=\{1,\dots,m\}$, and $AV=VD$.

We apply the approach in terms of Equation~(\ref{eq:state-perf}), under the assumptions; $\dot{\mathbf{q}}=A\mathbf{q}$. Initial guess is expressed as a combination of eigenvectors $\mathbf{q}_0=\mathbf{\gamma}V$. The generic solution is $\mathbf{q}=\mathbf{\gamma}e^{Dt}V$. % to rewrite this so it's exactly as in the proof

Let us consider the linear combination of all the coefficients of the state at time $t$. This can be expressed in the following form
\begin{equation}\label{eq:proof-output}\begin{split}
  \gamma_0e^{\lambda_0t}v_0/T+&(1/T)\sum_{j=1}^r{\gamma_je^{\lambda_j^0t}v_j^0}+\\&(1/T)\sum_{j=1}^r{\gamma_je^{\lambda_j^1t}v_j^1},
\end{split}\end{equation}
we proof that the eigenvalues $\mathbf{\lambda}$ and eigenvectors $V$ are such that Equation~(\ref{eq:proof-output}) is equivalent to Equation~(\ref{eq:proof-complex}). % the V and gamma are wrong sizes. You put zeros so you fill them (look at the sheet where you sketched it before)

The matrix $A$ is a block diagonal matrix, so we can express its determinant as the multiplication of the determinants of its blocks $\det{(A)}=\det{(0)}\times\det{(A_1)}\times\cdots\times\det{(A_r)}$. We proof the first determinant and the others separately.

We proof that the first term of both the Equation~(\ref{eq:proof-complex}) and~(\ref{eq:proof-output}) matches. We find the eigenvalue from $\det(0)=0$, which is $\lambda_0=0$. The corresponding eigenvector can be chosen arbitrarily $(0-\lambda_0)v_0=0\,\,\,\forall v_0$, thus we choose $v_0=1$. We find the value $\gamma_0$ of the vector $\gamma$ so that the terms are equal $a_0=\gamma_0$.

For the first block $A_1$, we find the eigenvalues from $\det(A_1-\lambda I)=0$. The polynomial $\lambda^2+\omega^2$, gives two complex roots--the two eigenvalues $\lambda_1^0=i\omega$ and $\lambda_1^1=-i\omega$. The eigenvector associated with the eigenvalue $\lambda_1^0$ is $v_1^0=\begin{bmatrix}-i&1\end{bmatrix}^T$. The eigenvector associated with the eigenvalue $\lambda_1^1$ is $v_1^1=\begin{bmatrix}i&1\end{bmatrix}^T$. Again, we find the values $\gamma_1$ of the vector $\gamma$ such that the equivalences $e^{i\omega t}(a_1-ib_1)=\gamma_1e^{i\omega t}\begin{bmatrix}-i&1\end{bmatrix}^T$ and $e^{-i\omega t}(a_1+ib_1)=\gamma_1e^{i\omega t}\begin{bmatrix}i&1\end{bmatrix}^T$ hold. They hold for $\gamma_1=\begin{bmatrix}b_1&a_1\end{bmatrix}$.

We thus conclude that the lemma holds.

\qed

\section{Mission Plan Format}
